
---
output: 
  github_document:
    html_preview: true
---

<!-- README.md is generated from README.Rmd. Please edit README.Rmd (this file) -->


```{r globaloptions, echo = FALSE}
knitr::opts_chunk$set(fig.path='Figs/',
                      fig.align = "left",
                      echo=FALSE, warning=FALSE, message=FALSE)
```

# R-MERMAID
A repo of useful R code for MERMAID analyses and figures. 

Contains a dataset of collected underwater observations from WCS and Manta Trust coral bleaching surveys from the Yasawas, Ra, Ovalau and Leleuvia in Fiji collected between April-June 2019. Learn more about the rapid coral bleaching assessment (http://www.wcs.org/coral)

We  plan to add more resource code to support MERMAID analysis in the future. 

Exciting demo showing how the data and code produces a histogram of coral bleaching and a map of coral bleaching reports from the dataset. All code draws on a real dataset of bleaching observations (by coral genus) downloaded from MERMAID (http://www.datamermaid.org)

```{r load script, results = "hide", include = FALSE}
library(here)
source(here::here("R", "bleaching-surveys", "02-bleach-summaries.R"))
source(here::here("R", "bleaching-surveys", "03-bleaching-histograms.R"))
source(here::here("R", "bleaching-surveys", "04-fiji-bleaching-map.R"))
source(here::here("R", "bleaching-surveys", "05-fiji-bleaching-month.R"))
```

```{r bleachinghistogram, fig.width=6, fig.height = 2.5, fig.align="center"}
ggplot(data = data,
       aes(x = floor(data$perc_bleach))) + 
  geom_histogram(aes(fill = bleach.class), 
                 colour = "black", binwidth = 2) + 
  theme_sleek(base_size = 14) +
  scale_fill_manual(values = c("olivedrab3","yellow","orange","red"), 
                    "Bleached colonies", 
                    labels = c("<4%", "5-20%", "20-40%", ">40%")) +
  xlab("Percent bleached colonies") +
  ylab("# of reefs") +
  scale_x_continuous(breaks = c(0,10,20,30,40,50,60,70,80), expand = c(0.05,0.5)) + 
  scale_y_continuous(expand = c(0,0), 
                     limits = c(0,15))
```


```{r bleachingmap, fig.width=10, fig.height=4, fig.retina=2}
ggplot() +
  geom_polygon(data = fiji, 
               aes(x= long, y= lat, group = group)) +
  coord_equal() + 
  scale_x_continuous(limits = c(176.5,181.5)) + 
  geom_point(data = data,
             aes(x = longitude, y = latitude, fill = bleach.class), 
             size = 5, shape = 21, alpha = 0.5, 
             colour = "black", 
             position = position_jitter(width = 0.005, height = 0.005)) + 
  scale_fill_manual(values = c("olivedrab3","yellow","orange","red"), 
                    "Bleached colonies", 
                    labels = c("<4%", "5-20%", "20-40%", ">40%")) +
  theme_sleek(base_size = 18) + 
  #theme(legend.position = "none")
  guides(fill = guide_legend(override.aes = list(alpha = 1))) + 
  xlab("Longitude") + 
  ylab("Latitude")
```

```{r bleaching-map-month, fig.width=10, fig.height=3, fig.retina=2}
ggplot() +
  geom_polygon(data = fiji, 
               aes(long, lat, group = group)) +
  coord_equal() + 
  scale_x_continuous(limits = c(176.5,181.5)) + 
  geom_point(data = data,
             aes(x = longitude, y = latitude, fill = bleach.class), 
             size = 5, shape = 21, alpha = 0.5, 
             colour = "black", 
             position = position_jitter(width = 0.05, height = 0.05)) + 
  scale_fill_manual(values = c("olivedrab3","yellow","orange","red"), 
                    "Bleached colonies", 
                    labels = c("<4%", "5-20%", "20-40%", ">40%")) +
  theme_sleek(base_size = 16) + 
  #theme(legend.position = "none") + 
  guides(fill = guide_legend(override.aes = list(alpha = 1))) + 
  xlab("Longitude") + 
  ylab("Latitude") + 
  facet_wrap(~month, nrow = 1)
```

With thanks to data contributors: Stacy Jupiter (WCS Melanesia), Sangeeta Mangubhai (WCS Fiji), Emily Darling (WCS Global) and Rob Macfarlane (Barefoot Manta Resort, Fiji)